Analysing Constraint Grammars with a SAT-solver

### 2016-02-29 paper/lrec/problem_desc.tex
more text in intro & add label for conclusions

\begin{itemize}
\item Are all the rules distinct? (e.g. \texttt{Para} and \texttt{De} may be included in \texttt{Prep})
\item Could two or more rules be merged? (e.g. \texttt{SELECT Inf IF -1 Prep OR Vai OR Vbmod ...})
\item What is the best order for the rules?
\item Generate a sequence that triggers rule(s) $R$ but not rule(s) $R'$. 
\end{itemize}

@@

For the second use case, here are examples of conflicts that our tool will detect.
\begin{itemize}
\item If a rule appears twice, the second occurrence will be disabled by the first
\item $R$ selects something in a context, $r$ removes~it
\item $R$ removes something from the context of $r$, so $r$ can never
  apply
\item $r$ has an internal conflict, such as non-existent
tag combination, or contradictory requirements for a context word
\end{itemize}
$R$ can be one rule or a set of rules: for instance, if one rule removes a verb in
context $C$, and another in context $\neg C$, together these rules
remove a verb in all possible cases, disabling any future rule that
targets verbs.

While rule-internal conflicts can be detected by simpler means, taking
care of rule interaction requires a {\em semantic} rather than a {\em
 syntactic} analysis.
% In order to find effects of rule interaction, we cannot just apply the
% rules to some actual sentences and hope to find a conflict---corpora
% can be extremely valuable in the development of CG rules, but there
% are languages where corpora are not available, or they may contain
% text from limited domains. 
We must keep track of all the possible sentences after applying each
rule. At each step, we have two options: either the rule fires,
or it does not fire. In case the rule does not fire, we have again two
options: either its conditions are not met,
or its target is the only remaining analysis. This results in a complex table
of interdependent decisions---some determined by being the target of a
rule, others by being a condition.

We express these constraints as a \emph{Boolean satisfiability problem} (SAT).
SAT problems consist of two components: a set of Boolean variables, and a set
of clauses on those variables. For instance, let the set be $\{a,b\}$
and the formulas $\{a \vee b, \neg{}a\}$. A program called \emph{SAT
  solver} will try to find a solution, where all the variables have a
value. For this particular problem, the unique solution is $a=False,
b=True$. It is also possible for a SAT problem to have no solution, or
multiple solutions.

@@

%The paper is structured as follows. Section~\ref{sec:prev} relates our work to the previous research. 
% Section~\ref{sec:implementation} discusses the implementation, 
% Section~\ref{sec:eval} presents preliminary results and discusses future work. 
%Section~\ref{sec:conclusion} concludes the paper.

>>>>

\begin{itemize}
\item Are all the rules distinct? (e.g. \texttt{Para} and \texttt{De} may be included in \texttt{Prep})
\item Can two or more rules be merged? (e.g. \texttt{SELECT Inf IF (-1C Prep OR Vai OR Vbmod ...)})
\item What is the best order for the rules?
\item Generate a sequence that triggers rules $R$ but not rules $R'$. 
\end{itemize}

@@

For the second use case, here are examples of conflicts that our tool will detect.
\begin{itemize}
\item If a rule appears twice, the second occurrence will be disabled by the first
\item $r$ selects something in a context, $r'$ removes~it
\item $R$ removes something from the context of $r$, so $r$ can never
  apply
\item $r$ has an internal conflict, such as non-existent
tag combination, or contradictory requirements for a context word
\end{itemize}
$R$ can be a single rule or a set of rules: for instance, if one rule removes a verb in
context $C$, and another in context $\neg C$, together these rules
remove a verb in all possible cases, disabling any future rule that
targets verbs.

While rule-internal conflicts can be detected by simpler means, taking
care of rule interaction requires a {\em semantic} rather than a {\em
 syntactic} analysis.
% In order to find effects of rule interaction, we cannot just apply the
% rules to some actual sentences and hope to find a conflict---corpora
% can be extremely valuable in the development of CG rules, but there
% are languages where corpora are not available, or they may contain
% text from limited domains. 
We must keep track of all the possible sentences after applying each
rule. At each step, we have two options: either the rule fires,
or it does not fire. In case the rule does not fire, we have two
reasons why not: one or more of its conditions does not hold, or its
target is the only remaining analysis. The result is a complex table
of interdependent decisions---some determined by being the target of a
rule, others by being a condition.

We express these constraints as a \emph{Boolean satisfiability problem} (SAT).
A SAT-problem consists of two components: a set of Boolean variables, and a set
of clauses on those variables. For instance, let the set ov variables
be $\{a, b\}$ and the formulas $\{a \vee b, \neg{}a\}$. A program called
\emph{SAT-solver} will try to find a solution, where all the variables
are replaced by a Boolean value. For this particular problem, the
unique solution is $\{a=False, b=True\}$, but it is also possible for a
SAT-problem to have no solution or multiple solutions.
In the case of CG analysis, the model we build is a \emph{sentence},
which starts as potentially anything, and is being shaped by all the
rules along the way. This is how we can generate possible inputs, as
well as check if the grammar is internally consistent.

The paper is structured as follows. Section~\ref{sec:prev} relates our
work to the previous research. 
Section~\ref{sec:implementation} discusses the implementation more in depth, and
Section~\ref{sec:eval} presents preliminary results. 
Section~\ref{sec:conclusion} discusses future work and concludes the paper.

@@

% Section~\ref{sec:implementation} discusses the implementation, 
% Section~\ref{sec:eval} presents preliminary results and discusses future work. 
%Section~\ref{sec:conclusion} concludes the paper.
